paper3

### 2022-07-22 paper3.Rmd
Bayesian ranking

## Download the wildness measure and its components
#"https://github.com/yiminchangchien/scenicness/blob/2a3f9b996c11a7c480867524664401dbea9aea2b/Wildness.RData" %>% url() %>%

load("C:\\Users\\00631\\Downloads\\James\\Wildness.RData")

## Download Scenic-Or-Not data
sc <- read_tsv("http://scenicornot.datasciencelab.co.uk/votes.tsv",
               col_types = cols("ID" = col_number(),
                                "Lat" = col_double(),
                                "Lon" = col_double(),
                                "Average" = col_double(),
                                "Variance" = col_double(),
                                "Votes" = col_character(),
                                "Geograph URI" = col_character())) %>%
  rowwise() %>%
  mutate(Mean = Votes %>% strsplit(",") %>% unlist() %>% as.numeric() %>% mean(),
         Variance = Votes %>% strsplit(",") %>% unlist() %>% as.numeric() %>% var(),
         Entropy = Votes %>% strsplit(",") %>% unlist() %>% as.numeric() %>% entropy(),
         Number = Votes %>% strsplit(",") %>% unlist() %>% as.numeric() %>% na.omit() %>% length())%>%
  st_as_sf(coords = c("Lon","Lat"), crs = 4326) %>%
  st_transform(crs = 27700)
```

```{r}
LANDMAP <- sc %>% 
  st_join(VS, .) %>% 
  aggregate(Votes~UID, data = ., paste0, collapse = ",") %>%
  merge(VS, ., by = "UID", all = T) %>%
  as_tibble() %>% 
  st_as_sf() %>%
  rowwise() %>%
  mutate(Scenicness.median = Votes %>% strsplit(",") %>% unlist() %>% as.numeric() %>% median(),
         Scenicness.mean = Votes %>% strsplit(",") %>% unlist() %>% as.numeric() %>% mean(),
         Scenicness.entropy = Votes %>% strsplit(",") %>% unlist() %>% as.numeric() %>% entropy(),
         Scenicness.number = Votes %>% strsplit(",") %>% unlist() %>% as.numeric() %>% na.omit() %>% length(),
    Consultant = ifelse(str_detect(VS_1a, "Bronwen Thomas"), "B", "A")) %>%
  dplyr::select(UID, CLS_1, CLS_2, CLS_3, CLS_4, ScenicQuality, Consultant, Votes, Scenicness.median, Scenicness.mean, Scenicness.entropy, Scenicness.number) %>%
  ungroup()

@@

LANDMAP <- st_transform(LANDMAP, crs = st_crs(Wildness)) %>%
  exact_extract(Wildness, ., fun = c('mean', 'median', 'coefficient_of_variation', 'variance', 'quantile'), quantiles = c(0.25, 0.75)) %>%
  mutate(Wildness.iqr = q75 - q25) %>%
  dplyr::select(Wildness.mean = mean, Wildness.median = median, Wildness.iqr, Wildness.cv = coefficient_of_variation, Wildness.variance = variance) %>%
  bind_cols(LANDMAP, .) %>%
  st_transform(crs = 27700)

landmap <- st_cast(LANDMAP, "POLYGON")
```

```{r}
library(reticulate)

```{python}

import geopandas as gpd
import pandas as pd
from sklearn.preprocessing import OrdinalEncoder
from sklearn.compose import make_column_transformer

# landmap = geopandas.read_file("/Users/mac/R/LCA-GB/Data/Input/LANDMAP.geojson")
landmap = r.landmap
landmap = landmap.dropna(subset=['ScenicQuality'])

print(landmap)
print(landmap.describe)
print(landmap['ScenicQuality'].value_counts())

@@

# split train/test set
X = landmap.loc[:, ['Scenicness.mean']]
y = landmap.loc[:, ['ScenicQuality']]

@@

scale_mapper = {'Low': int(0), 'Moderate': int(1), 'High': int(2), 'Outstanding': int(3)} 
y[['ScenicQuality']] = y[['ScenicQuality']].replace(scale_mapper).astype('int')

@@

from sklearn.model_selection import train_test_split
# Create train/test split
X_train, X_test, y_train, y_test = train_test_split(X, y, test_size=0.2, random_state=2046, stratify=y)

## Hyperparameter Tuning

@@

The domain space is the input values over which we want to search.

Choose hyperparameter domain to search over

@@

```{python}
# -- hyperopt for optimising a 
from hyperopt import fmin, tpe, hp, STATUS_OK, Trials, space_eval
from sklearn.metrics import cohen_kappa_score
import numpy as np
import xgboost as xgb

@@

  # Booster Parameters      
               'eta': hp.uniform('eta', 0.01, 1.01),
             'gamma': hp.quniform('gamma', 0.1, 5, 0.05), 
         'max_depth': hp.choice('max_depth', np.arange(1, 30, 1, dtype=int)),
  'min_child_weight': hp.choice('min_child_weight', np.arange(1, 30, 1, dtype=int)),     
         'subsample': hp.quniform('subsample', 0.3, 1.01, 0.05),
  'colsample_bytree': hp.uniform('colsample_bytree', 0.5, 1.0),
        'reg_lambda': hp.uniform('reg_lambda', 0.0, 1.0),
         'reg_alpha': hp.quniform('reg_alpha', 40, 180, 1),
  #      'tree_method': 'hist','gpu_hist',  # Use GPU accelerated algorithm
  # 'scale_pos_weight':
  #       'max_leaves':
  #     'n_estimators': hp.quniform('n_estimators', 5, 35, 1),
  #       'num_leaves': hp.quniform('num_leaves', 5, 50, 1),
  #        'predictor': 'cpu_predictor'

  # Learning Task Parameters
         'num_class': 4,                   # Number of possible output classes
         'objective': 'multi:softmax',     # Specify multiclass classification
       'eval_metric': 'merror',
              'seed': 2046
    }

# Sample values from these configuration spaces using the routines in hyperopt.pyll.stochastic
# from hyperopt.pyll.stochastic import sample
# space = {'learning_rate': hp.loguniform('learning_rate', np.log(0.01), np.log(1))}
# print(sample(space))
```

#### 2. Define an objective/loss/cost function

@@

```{python}
# -- optimize it using algorithm of choice
def optimize(trials, space):
    best = fmin(fn = objective,
                space = param_space, 
                algo = tpe.suggest,    # the optimisation algorithm
                max_evals = 10,        # the number of iteration
                trials = trials)       # the trials database for fmin
    return best

@@

# Return the best parameters
space_eval(param_space, best_params)

# -- print out what we got!
# for trial in trials.trials:
#     print(trial)
```
#### 4. Results

@@

```{python}
import xgboost as xgb
# Specify sufficient boosting iterations to reach a minimum
num_round = 1000 #209.91

# Convert input data from numpy to XGBoost format
dtrain = xgb.DMatrix(X_train, label = y_train)
 dtest = xgb.DMatrix(X_test, label = y_test)

import time
gpu_res = {}  # Store accuracy result
start = time.time()

# Train model
# model = xgb.train(param, dtrain, num_round, evals=[(dtest, 'test')], evals_result=gpu_res)
# print("GPU Training Time: %s seconds" % (str(time.time() - start)))

best_params['num_class'] = 4
best_params['objective'] = 'multi:softmax'
best_params['eval_metric'] = 'merror'

model = xgb.train(best_params,
                  dtrain,
                  num_round = 1000)
                  #evals = [(dtrain, 'train'),(dtest, 'test')])

print(time.time() - start)

@@

# Evaluate model using test data
from sklearn.metrics import accuracy_score

@@

from sklearn.metrics import cohen_kappa_score
kappa = cohen_kappa_score(y_test, y_pred)
print("Kappa: %.2f%%" % (kappa))

# save to JSON
# model.save_model("/Users/mac/PycharmProjects/XGBoost/models/model_A.json")
```

@@

```{python}
explainer_shap = shap.TreeExplainer(final_model)
shap_values = explainer_shap(X_coords)

shap_interaction_values = shap.TreeExplainer(
    final_model).shap_interaction_values(X_coords)

X_names = ['pct_18_34','pct_white','pct_bach',
          'pct_no_car','log_pop_den','job_entropy','network_den',
          'ave_trip_dist','pct_share','x_coord','y_coord']

fig, ax = plt.subplots(figsize=(8, 6),dpi=160)

shap.summary_plot(shap_interaction_values, X_coords, max_display=15, 
                  feature_names = X_names,
                  plot_type="compact_dot")
```

>>>>

## Download Scenic-Or-Not data
sc1 <- read_tsv("http://scenicornot.datasciencelab.co.uk/votes.tsv",
               col_types = cols("ID" = col_number(),
                                "Lat" = col_double(),
                                "Lon" = col_double(),
                                "Average" = col_double(),
                                "Variance" = col_double(),
                                "Votes" = col_character(),
                                "Geograph URI" = col_character())) %>%
  rowwise() %>%
  mutate(Mean = Votes %>% strsplit(",") %>% unlist() %>% as.numeric() %>% mean(),
         Median = Votes %>% strsplit(",") %>% unlist() %>% as.numeric() %>% median(),
         Variance = Votes %>% strsplit(",") %>% unlist() %>% as.numeric() %>% var(),
         Entropy = Votes %>% strsplit(",") %>% unlist() %>% as.numeric() %>% entropy(),
         Number = Votes %>% strsplit(",") %>% unlist() %>% as.numeric() %>% length(),
           `10` = Votes %>% strsplit(",") %>% unlist() %>% str_count("10") %>% sum(),
            `9` = Votes %>% strsplit(",") %>% unlist() %>% str_count("9") %>% sum(),
            `8` = Votes %>% strsplit(",") %>% unlist() %>% str_count("8") %>% sum(),
            `7` = Votes %>% strsplit(",") %>% unlist() %>% str_count("7") %>% sum(),
            `6` = Votes %>% strsplit(",") %>% unlist() %>% str_count("6") %>% sum(),
            `5` = Votes %>% strsplit(",") %>% unlist() %>% str_count("5") %>% sum(),
            `4` = Votes %>% strsplit(",") %>% unlist() %>% str_count("4") %>% sum(),
            `3` = Votes %>% strsplit(",") %>% unlist() %>% str_count("3") %>% sum(),
            `2` = Votes %>% strsplit(",") %>% unlist() %>% str_count("2") %>% sum(),
            `1` = Votes %>% strsplit(",") %>% unlist() %>% str_count("1") %>% sum())%>%
  mutate(`1` = `1`-`10`) %>%
  st_as_sf(coords = c("Lon","Lat"), crs = 4326) %>%
  st_transform(crs = 27700)
```

```{python}
import math
from scipy import stats

def ci_lower_bound(positive_ratings, total_ratings, confidence):
    """
    Following http://www.evanmiller.org/how-not-to-sort-by-average-rating.html#changes
    :return: the lower bound of the confidence interval
    """
    n =  total_ratings
    z = stats.norm.ppf(1 - (1 - confidence) / 2)
    p_hat = 1.0 * positive_ratings / n
    result = (
        (
             p_hat + z * z / (2 * n)
             - z * math.sqrt(
                 (p_hat * (1 - p_hat) + z * z / (4 * n)) / n
             )
        ) / (1 + z * z / n)
    )
    return result

```{r}
LANDMAP <- sc %>% 
  st_join(VS, .) %>% 
  aggregate(Votes~UID, data = ., paste0, collapse = ",") %>%
  merge(VS, ., by = "UID", all = T) %>%
  as_tibble() %>% 
  st_as_sf() %>%
  rowwise() %>%
  mutate(Scenicness.mean_all = Votes %>% strsplit(",") %>% unlist() %>% as.numeric() %>% mean(),
         Scenicness.median_all = Votes %>% strsplit(",") %>% unlist() %>% as.numeric() %>% median(),
         Scenicness.entropy_all = Votes %>% strsplit(",") %>% unlist() %>% as.numeric() %>% entropy(),
         Scenicness.number_all = Votes %>% strsplit(",") %>% unlist() %>% as.numeric() %>% na.omit() %>% length()) %>%
  ungroup()

LANDMAP <- sc %>%
  st_join(LANDMAP, .) %>%
  aggregate(Median~UID, data = ., paste0, collapse = ",") %>%
  merge(LANDMAP, ., by = "UID", all = T) %>%
  as_tibble() %>%
  st_as_sf() %>%
  rowwise() %>%
  mutate(Scenicness.mean_of_median = Median %>% strsplit(",") %>% unlist() %>% as.numeric() %>% mean(),
         Scenicness.median_of_median = Median %>% strsplit(",") %>% unlist() %>% as.numeric() %>% median()) %>%
  ungroup()

LANDMAP <- sc %>%
  st_join(LANDMAP, .) %>%
  aggregate(Mean~UID, data = ., paste0, collapse = ",") %>%
  merge(LANDMAP, ., by = "UID", all = T) %>%
  as_tibble() %>%
  st_as_sf() %>%
  rowwise() %>%
  mutate(Scenicness.mean_of_mean = Mean %>% strsplit(",") %>% unlist() %>% as.numeric() %>% mean(),
         Scenicness.median_of_mean = Mean %>% strsplit(",") %>% unlist() %>% as.numeric() %>% median(),
         Consultant = ifelse(str_detect(VS_1a, "Bronwen Thomas"), "B", "A")) %>%
  dplyr::select(UID, CLS_1, CLS_2, CLS_3, CLS_4, ScenicQuality, Consultant, 
                Votes, Scenicness.mean_all, Scenicness.median_all, Scenicness.entropy_all, Scenicness.number_all,
                Scenicness.mean_of_median, Scenicness.median_of_median,
                Scenicness.mean_of_mean, Scenicness.median_of_mean) %>%
  ungroup()

@@

LANDMAP <- st_transform(LANDMAP, crs = st_crs(Wildness)) %>%
  exact_extract(Wildness, ., fun = c('mean', 'median', 'coefficient_of_variation', 'variance', 'quantile'), quantiles = c(0.25, 0.75)) %>%
  mutate(Wildness.iqr = q75 - q25) %>%
  dplyr::select(Wildness.mean = mean, Wildness.median = median, Wildness.iqr, Wildness.cv = coefficient_of_variation, Wildness.variance = variance) %>%
  bind_cols(LANDMAP[, 1:13], .) %>%
  st_transform(crs = 27700)

landmap <- st_cast(LANDMAP, "POLYGON") %>%
  dplyr::mutate(x = st_centroid(landmap$geometry) %>% st_coordinates() %>% .[, 1],
                y = st_centroid(landmap$geometry) %>% st_coordinates() %>% .[, 2])
summary(landmap)
```

```{r}
Sys.setenv("RETICULATE_PYTHON" = ".\\.venv\\Scripts\\python.exe")
# Sys.getenv()
Sys.which("Python")

# py_discover_config()

library(reticulate)
# -- creates an interactive Python console, REPL (Read, Eval, Print Loop), within R.
repl_python()

# reticulate::py_config() # why numpy_version couldn't be found?
# use_python("Z:\\R\\R projects\\LCA-GB\\my_env_Python_3.9.13\\Scripts\\python.exe")
# 
# # use_virtualenv("Z:\\R\\R projects\\LCA-GB\\my_env_Python_3.9.13", required = TRUE)
# 
# install.packages("usethis")
# usethis::edit_r_environ()
# usethis::edit_r_profile()
```

```{python}
import geopandas as gpd
import pandas as pd
from sklearn.preprocessing import OrdinalEncoder
from sklearn.compose import make_column_transformer

landmap = r.landmap
landmap = landmap.dropna(subset=['ScenicQuality'])

print(landmap)
landmap.describe()
print(landmap['ScenicQuality'].value_counts())

@@

# split train/test set
# X = landmap_B.loc[:, ['Scenicness.median', 'Scenicness.entropy', 
#                       'Wildness.median', 'Wildness.variance',
#                       'x', 'y']]

X = landmap_B.loc[:, ['Scenicness.mean', 'Wildness.mean']]
y = landmap_B.loc[:, ['ScenicQuality']]

@@

scale_mapper = {'Low': int(0), 
                'Moderate': int(1), 
                'High': int(2), 
                'Outstanding': int(3)}

y[['ScenicQuality']] = y[['ScenicQuality']].replace(scale_mapper).astype('int')

@@

# -- Create train/test split
from sklearn.model_selection import train_test_split
X_train, X_test, y_train, y_test = train_test_split(X, y, 
                                                    test_size = 0.2, 
                                                    random_state = 2046, 
                                                    stratify = y)
```
## Hyperparameter Tuning

@@

The domain space is the input values over which we want to search.
Choose hyperparameter domain to search over

@@

```{python}
# -- hyperopt for optimising a XGBoost model
from hyperopt import fmin, tpe, hp, STATUS_OK, Trials, space_eval
from sklearn.metrics import cohen_kappa_score
import numpy as np
import xgboost as xgb

@@

  # Booster Parameters
   'num_boost_round': hp.quniform('num_boost_round', 100, 1500, 0.001), 
               'eta': hp.uniform('eta', 0.01, 1.01),
             'gamma': 0, 
         'max_depth': hp.choice('max_depth', np.arange(1, 10, 1, dtype=int)),
  'min_child_weight': hp.quniform('min_child_weight', 0, 1, 0.1),     
         'subsample': hp.quniform('subsample', 0.3, 1, 0.001),
  'colsample_bytree': 1, #hp.uniform('colsample_bytree', 0.5, 1.01),
 #       'reg_lambda': hp.uniform('reg_lambda', 0.0, 1.0),
 #        'reg_alpha': hp.quniform('reg_alpha', 40, 180, 1),
 #      'tree_method': 'hist','gpu_hist',  # Use GPU accelerated algorithm
 # 'scale_pos_weight':
 #       'max_leaves':
 #     'n_estimators': hp.quniform('n_estimators', 5, 35, 1),
 #       'num_leaves': hp.quniform('num_leaves', 5, 50, 1),
 #        'predictor': 'cpu_predictor',

  # Learning Task Parameters
         'num_class': 4,                   # Number of possible output classes
         'objective': 'multi:softmax',     # Specify multi-class classification
       'eval_metric': 'merror',
              'seed': 2022
    }

# Sample values from these configuration spaces using the routines in hyperopt.pyll.stochastic
from hyperopt.pyll.stochastic import sample
space = {'eta': hp.uniform('eta', 0.01, 1.01)}
print(sample(space))
```
#### 2. Define an objective/loss/cost function

@@

```{python}
# -- optimize it using algorithm of choice
def optimize(trials, space):
    best = fmin(fn = objective,
                space = param_space, 
                algo = tpe.suggest,    # the optimisation algorithm
                max_evals = 30,        # the number of iteration
                trials = trials)       # the trials database for fmin
    return best

@@

# Return the best parameters
space_eval(param_space, best_params)
```
#### 4. Results

@@

```{python}
import xgboost as xgb
## Specify sufficient boosting iterations to reach a minimum
## num_round = 1000

# -- convert input data from numpy to XGBoost format
dtrain = xgb.DMatrix(X_train, label = y_train)
 dtest = xgb.DMatrix(X_test, label = y_test)

# -- try previous combinations of hyperparameter set
# Consultant A
# param_space = {
#            'booster': 'gbtree',
#    'num_boost_round': 209.910,
#                'eta': 0.451,
#              'gamma': 0,
#          'max_depth': 12,
#   'min_child_weight': 0.2,
#          'subsample': 0.758,
#   'colsample_bytree': 1,
#          'num_class': 4,
#          'objective': 'multi:softmax',
#        'eval_metric': 'merror',
#               'seed': 2046
#     }

# Consultant B
params = {
          'booster': 'gbtree',
        #'num_round': 1433.996,
              'eta': 0.892,
            'gamma': 0,
        'max_depth': 4,
 'min_child_weight': 0.6,
        'subsample': 0.500,
 'colsample_bytree': 1,
        'num_class': 4,
        'objective': 'multi:softmax',
      'eval_metric': 'merror',
             'seed': 1997
  }

# -- train model
import time
start = time.time()

model = xgb.train(space_eval(param_space, best_params), 
                  dtrain, 
                  num_boost_round = 500,
                  evals = [(dtrain, 'train'), (dtest, 'test')],
                  early_stopping_rounds = 250)

# model = xgb.train(params, 
#                   dtrain, 
#                   num_boost_round = 500, 
#                   evals = [(dtrain, 'train'), (dtest, 'test')],
#                   early_stopping_rounds = 250)

# model = xgb.cv(space_eval(param_space, best_params), dtrain, num_boost_round = 2000, nfold = 5, stratified = True)
# model.head()

print("XGBoost model training time: %s seconds" % (str(time.time() - start)))

@@

# -- evaluate model using test data
from sklearn.metrics import accuracy_score

@@

from sklearn.metrics import cohen_kappa_score
kappa = cohen_kappa_score(y_pred, y_test)
print("Kappa: %.2f" % (kappa))

# -- save to JSON
model.save_model("Z:\\R\\R projects\\LCA-GB\\models\\model_B.json")
```

@@

```{python}
import shap
import numpy as np
import matplotlib.pyplot as plt
from sklearn.preprocessing import StandardScaler
from sklearn.metrics import plot_confusion_matrix

# -- load the saved model
model = xgb.XGBRegressor()
model.load_model("Z:\\R\\R projects\\LCA-GB\\models\\model_B.json")

# -- view the confusion matrix
class_names = ['Low', 'Moderate', 'High', 'Outstanding']
disp = plot_confusion_matrix(model, X_test, y_test, 
                             display_labels = class_names, 
                             cmap = plt.cm.Blues, 
                             xticks_rotation = 'vertical')

# -- view the feature importance
importances = model.feature_importances_
indices = np.argsort(importances)
features = X.columns

fig, ax = plt.subplots()

plt.title('Feature Importances')
plt.barh(range(len(indices)), importances[indices], color='g', align='center')
plt.yticks(range(len(indices)), [features[i] for i in indices])
plt.xlabel('Relative Importance')
plt.show()

fig.set_size_inches(18.5, 10.5)
fig.savefig("Z:\\R\\R projects\\LCA-GB\\models\\pyplot.png")
plt.clf()

# -- play shap
explainer = shap.TreeExplainer(model)
shap_values = explainer.shap_values(X_test, approximate = True)
shap.summary_plot(shap_values[0], X_test.values, feature_names = X_test.columns)
fig.set_size_inches(18.5, 20.5)
fig.savefig("Z:\\R\\R projects\\LCA-GB\\models\\shap.png")

# -- first trial
shap.initjs()
shap_values = shap.TreeExplainer(model).shap_values(X_test)
shap.summary_plot(shap_values, X_test)

# -- second trial
scaler = StandardScaler()
scaled_data = scaler.fit_transform(X_test.iloc[:,:-1])
shap_values = explainer.shap_values(scaled_data.iloc[0,:-1].values.reshape(1,-1))
shap.force_plot(explainer.expected_value[0], shap_values[0], link='logit')  # repeat changing 0 for i in range(0, 5)

# -- third trial
shap.initjs()
class_names = ['Low', 'Moderate', ' High', 'Outstanding']
shap.summary_plot(shap_values, X_test, plot_type="bar", class_names= class_names, feature_names = ['Scenicness', 'Wildness'])

# -- plot the global importance of each feature
shap.plots.bar(explainer)

# visualize the first prediction's explanation
shap.force_plot(explainer.expected_value[0], shap_values[0])

# fig, ax = plt.subplots(figsize=(20, 20),dpi=200)
# shap.summary_plot(shap_values)
# # or 
# shap.plots.beeswarm(shap_values)
shap.summary_plot(shap_values[2], X_test)

# fig, ax = plt.subplots(figsize=(20, 20),dpi=200)
shap.summary_plot(shap_values, X_test, plot_type="bar")

shap_interaction_values = shap.TreeExplainer(model).shap_interaction_values(X_train)

X_names = ['Scenicness.median','x_coord','y_coord']

shap.summary_plot(shap_interaction_values, X_train, max_display=15, 
                  feature_names = X_names,
                  plot_type = 'dot')
                  # plot_type="compact_dot")

import shap  # package used to calculate Shap values

# Create object that can calculate shap values
explainer = shap.TreeExplainer(my_model)

# calculate shap values. This is what we will plot.
# Calculate shap_values for all of val_X rather than a single row, to have more data for plot.
shap_values = explainer.shap_values(val_X)

# Make plot. Index of [1] is explained in text below.
shap.summary_plot(shap_values[1], val_X)                  
```
